|

### 2024-08-20 reproducibility/manuscript/manuscript.qmd
docs(manuscript): add cell fate uncertainty estimation methods
Signed-off-by: Cameron Smith <cameron.ray.smith@gmail.com>

>>>>

## Cell fate uncertainty estimation {#sec-methods-fate-uncertainty}

Pyro&thinsp;-Velocity evaluates cell fate uncertainty using four statistics. The
first is the uncertainty of shared time, which is evaluated using the standard
deviation of posterior samples of $t_c$ for each cell. The second statistic is
the angular standard deviation [@Berens2009-fu] across posterior samples of the
velocity-based vector field. The value of this statistic was mapped linearly to
the range $[0, 360]$ for interpretability. The third statistic is the Rayleigh
test [@The-Astropy-Collaboration2022-aw] for posterior samples of the
velocity-based vector fields for each cell, whose p-values have been corrected
for multiple tests using Benjamini-Hochberg FDR methods; if FDR is less than
$0.05$, the cell is considered to have one statistically significant future
direction. The second and third circular statistics were implemented using the
Astropy package (`v5.1`) [@The-Astropy-Collaboration2022-aw]. The last statistic
is state uncertainty which evaluates the mean Euclidean distance between
posterior mean and posterior samples of the raw read count prediction for each
cell
\begin{align*}
\frac{1}{N} \sum_{i} \sqrt{\sum_{g} \left(x_{icg} - \frac{1}{N}\sum_i x_{icg}\right)^2}.
\end{align*}
